---
title: 'Statistical report: Ablation in HF'
subtitle: 'DRAFT'
author: 'Statistician: Lina Benson'
  
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    number_sections: yes
    extra_dependencies: ["flafter"]
link-citations: yes
bibliography: references.bib
nocite: '@*'
urlcolor: blue
linkcolor: black
header-includes:
   - \usepackage{draftwatermark}
---

\newpage 
\tableofcontents 
\listoftables
\listoffigures
\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, include = TRUE, comment = "",
  warning = FALSE, message = FALSE, fig.pos = "H",
  out.extra = "",
  fig.path = "../output/figs/"
)
options(knitr.kable.NA = "")
```

```{r adjust_directory_if_needed, include=FALSE}
# Uncomment lines below if rmd file is placed in a subdirectory
knitr::opts_knit$set(root.dir = normalizePath("../"))
```

```{r load_project}
# 1. Set options in config/global.dcf
# 2. Load packages listed in config/global.dcf
# 3. Import functions and code in lib directory

ProjectTemplate::reload.project()

cacheon <- TRUE
```             

# Data handling

## Data source

SHFDB3, https://kiheartfailure.github.io/shfdb3/, v 3.2.4. 

## Inclusion/exclusion

```{r flowcase}
default_kable(flowcase, caption = "Flowchart for cases")
```

```{r flowcontrol}
default_kable(flowcontrol, caption = "Flowchart for controls")
```

The index date is defined as date of ablation for cases and index date (visit date/date of discharge) in SwedeHF for controls. 

First patient in: `r min(pop$indexdtm)` and last patient in: `r max(pop$indexdtm)`.  

The median age (IQR) is `r pop %>% summarise(med = fn(median(shf_age), dig = 1),
                                             q1 = fn(quantile(shf_age, probs = 0.25), dig = 1),
                                             q3 = fn(quantile(shf_age, probs = 0.75), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", q1, "-", q3, ")")) %>%
                                   pull(out)` and 
`r pop %>% count(shf_sex) %>%
  mutate(perc = fn(n / sum(n) * 100, 1)) %>%
  filter(shf_sex == "Female") %>%
  pull(perc)`% females.    

## Created variables 

```{r ddr}
default_kable(metalm, caption = "Medications from DDR")
```

```{r npr}
default_kable(metaout, caption = "Comorbidities and outcomes from Patient- and Cause of Death Registry")
```

\newpage

# Statistical analysis 

## General

All analyses were performed using `r sessionInfo()$R.version$version.string` [@r]. 
The level of significance is set to 5%, two-sided. No adjustment for multiple 
comparisons were made and therefore the results should be viewed with care.

## Missing data

The loss of patients due to missing data for the variables included in the models was small, `r paste0(pop %>% 
         mutate(anyna = factor(if_else(rowSums(is.na(select(., !!!syms(modvars)))) > 0, "Yes", "No"))) %>%
         group_by(ablation) %>%
         count(anyna, .drop = FALSE) %>%
         mutate(p = fn(n / sum(n) * 100, 1), 
                np = paste0(n, " (", p, "%)")) %>% 
         ungroup() %>%
         filter(anyna == "Yes") %>% 
         pull(np), collapse = ", ")` 
         for no ablation and ablation group respectively, and therefore no imputation was performed.

## Propensity scores and matching

A propensity score for ablation was estimated for each patient with 
logistic regression using the variables 
indicated in Table \ref{tab:tab1}, where `r paste0(nsvars, collapse = ", ")` 
were modelled using cubic splines with 4 degrees of freedom. 1:2 matching without replacement [@match] 
was thereafter performed, where matching was 
allowed if the propensity score differed by 0.01 or less. The ability of the 
propensity score matching to balance the baseline characteristics was assessed by 
standardized mean differences (SMD), where SMD |<0.1| was deemed 
acceptable to assume similarity between groups.  

1:2 has deemed the best option when taking into account the number of patients 
retained and balance between groups. Other matching ratios: `r matchingn`.

## Baseline characteristics

```{r, child = "../src/tab1.Rmd"}

```

## Assocations with outcomes

Data were censored at 2019-12-31, death, emigration, 5 years after index, or, for the controls, ablation for AF. 

Outcomes including all-cause mortality were presented with Kaplan-Meier curves and other outcomes 
with cumulative incidence curves treating death as a competing event. 

Cox proportional hazards regressions were 
used to model the time to event, partly crude and partly adjusted for variables indicated in 
Table \ref{tab:tab1}. The variables were selected based on clinical relevance. 
Adjustment was performed partly by including the variables as individual covariates directly 
in the model, where `r paste0(nsvars, collapse = ", ")` 
were modelled using cubic splines with 4 degrees of freedom, and partly using the propensity score matched cohort. 
In the latter a frailty term was included to model the dependence between matched pairs. 

As a consistency analysis, outcomes not including all-cause mortality were modelled using a sub-distributional hazards model [@fg] 
where death was treated as a competing event. 

The proportional hazards assumption was investigated using the scaled Schoenfeld 
residuals (cox.zph in [@survival-package]) and was for all outcomes deemed to be fulfilled for ablation. 

The median (min-max) follow-up was 
`r pop %>% summarise(med = fn(median(sos_outtime_death / 365.25), dig = 1),
                                             min = fn(min(sos_outtime_death / 365.25), dig = 1),
                                             max = fn(max(sos_outtime_death / 365.25), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", min, "-", max, ")")) %>%
                                   pull(out)` years for a total of 
                                   `r pop %>% summarise(sumpy = fn(sum(sos_outtime_death) / 365.25, dig = 0)) %>%
                                   pull(sumpy)` patient-years of follow-up.


```{r, child = "../src/km.Rmd"}

```

```{r, child = "../src/outtab.Rmd"}

```

\newpage

### EF subgroups

The association between ablation and the outcomes in EF subgroups were 
investigated using Cox proportional hazards regressions in the matched cohort 
where the interaction between ablation and EF was included, thereby giving the interpretation of a 
subgroup analysis but with the additional advantage of also being able to test 
for differences in effect between EF subgroups. 

```{r, child = "../src/outtabefsub.Rmd"}

```

### Residual confounding

In order to assess how robust the association between ablation and the primary outcome, all-cause-mortality/first HFH, is to residual confounding the E-value was calculated [@evalue]. The E-value is interpreted as the minimum Risk Ratio that an unmeasured confounder would need to have with both ablation and all-cause-mortality/first HFH, conditional on the variables currently adjusted for in the model, to fully eliminate the association between ablation and all-cause-mortality/first HFH.  

```{r, child = "../src/evalue.Rmd"}

```

See Table \ref{tab:evalue2} in order to compare the magnitude of the E-value within the context of the study [@evalue2]. 

\clearpage
\newpage

# Reproducibility

## R session information {#sessioninfo}

```{r sessinfo}
sessionInfo()
```

## R code

The R code for all data handling and statistical analyses are found: 
https://github.com/KIHeartFailure/ablation. On publication
the repository will be made public so as to 
link to it from the resulting article for increased transparency and code sharing.
No data or output is stored in the repository. 

# References
